---
title: "Visualization"
author: "Ducheng Tan"
date: "2026-05-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE) 
```

## Packages and Library Loadings 

```{r}
#install.packages("tidyverse")
#install.packages("palmerpenguins")
#install.packages("MASS")
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
library(ggplot2)
library(MASS)
library(ggplot2)
```

## Data and Introduction 

```{r}
palmerpenguins::penguins
glimpse(penguins)
```


```{r}
summary(penguins)
```
#ggplot 


```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = penguins$species)
) + geom_point(mapping = aes(color = species)) +  geom_smooth(method = "lm")
```
 
```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = species, shape =species)) +
  geom_smooth(method = "lm")
```
 
 
```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(aes(color = species, shape = species)) +
  geom_smooth(method = "lm") +
  labs(
    title = "Body mass and flipper length",
    subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
    x = "Flipper length (mm)", y = "Body mass (g)",
    color = "Species", shape = "Species"
  ) +
  scale_color_colorblind()
```
 
 
 1.2.5 Exercise 
 
How many rows are in penguins? How many columns?
344 x 8 (n x p) 

```{r}
nrow(penguins); ncol(penguins)
```


What does the bill_depth_mm variable in the penguins data frame describe? Read the help for ?penguins to find out.

a number denoting bill length (millimeters)  
```{r}
?penguins
```


Make a scatterplot of bill_depth_mm vs. bill_length_mm. That is, make a scatterplot with bill_depth_mm on the y-axis and bill_length_mm on the x-axis. Describe the relationship between these two variables.
There is a weak positive relation between the two variables. 

```{r}
ggplot(data = penguins, mapping = aes(x=bill_length_mm, y=bill_depth_mm, color = species)) + geom_point()
```


What happens if you make a scatterplot of species vs. bill_depth_mm? What might be a better choice of geom?
```{r}
ggplot(data = penguins, mapping= (aes(x= species,y= bill_depth_mm, color = species ))) +geom_boxplot()
```



Why does the following give an error and how would you fix it?
fixed via adding `mapping = aes(x, y)` 

```{r}
ggplot(data = penguins , mapping = aes(x= body_mass_g, y=flipper_length_mm, color =species)) +
  geom_point()
```



What does the na.rm argument do in geom_point()? What is the default value of the argument? Create a scatterplot where you successfully use this argument set to TRUE.

```{r}
ggplot(data = penguins , mapping = aes(x= body_mass_g, y=flipper_length_mm, color =bill_depth_mm)) +
  geom_point(na.rm = TRUE) # removes the missing data points without warning
```


Add the following caption to the plot you made in the previous exercise: “Data come from the palmerpenguins package.” Hint: Take a look at the documentation for labs().

```{r}
ggplot(data = penguins , mapping = aes(x= body_mass_g, y=flipper_length_mm, color =species)) +
  geom_point(na.rm = TRUE) +  labs(title = "body mass vs flipper_length",caption = "Data come from the palmerpenguins package.")
 # removes the missing data points without warning
```

Recreate the following visualization. What aesthetic should `bill_depth_mm` be mapped to? And should it be mapped at the global level or at the geom level?

```{r}
ggplot(data = penguins , mapping = aes(x=flipper_length_mm , y=body_mass_g, color =bill_depth_mm)) +
  geom_point(na.rm = TRUE, size = 3) + geom_smooth(method = "loess")
```


Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.
```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = island)
) +
  geom_point() +
  geom_smooth(se = FALSE)
```



Will these two graphs look different? Why/why not? 

```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point() +
  geom_smooth()

ggplot() +
  geom_point(
    data = penguins,
    mapping = aes(x = flipper_length_mm, y = body_mass_g)
  ) +
  geom_smooth(
    data = penguins,
    mapping = aes(x = flipper_length_mm, y = body_mass_g)
  )
```
 
##ggplot2 calls 



```{r}
 
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point()


```

The first two argument can be omitted after sometime and we can rewrite 

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point()
```


The pipe 
```{r}
penguins |> 
  ggplot(aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point()
```

## Visualizing Distributions 


# Categorical Variable 

```{r}
ggplot(penguins, aes(x = species, fill = species)) +geom_bar()
```

To further reorder the bars based on their frequencies we can transform the variables in a factor 

```{r}
ggplot(penguins, aes(x = fct_infreq(species), fill = species))

+geom_bar()
```


# Numerical variable 

Histograms for a continuous variable 
```{r}
ggplot(penguins, aes(x = body_mass_g)) +
  geom_histogram(binwidth = 200)
```
A histogram divides the x-axis into equally spaced bins and then uses the height of a bar to display the number of observations that fall in each bin. In the graph above, the tallest bar shows that 39 observations have a `body_mass_g value` between 3,500 and 3,700 grams, which are the left and right edges of the bar.

Too small or too big of the binwidth can hide the underlying distributions, use the correct binwidth to explain the data. 
For example: 
```{r}
ggplot(penguins, aes(x = body_mass_g)) +
  geom_histogram(binwidth = 20)
ggplot(penguins, aes(x = body_mass_g)) +
  geom_histogram(binwidth = 2000)
```
An alternative visualization for distributions of numerical variables is a density plot. A density plot is a smoothed-out version of a histogram and a practical alternative, particularly for continuous data that comes from an underlying smooth distribution. 

```{r, geom_density(emprical)}



ggplot(penguins, aes(x=body_mass_g)) + geom_density(linewidth =1, color = "dodgerblue")

```


