---
title: "Visualization"
author: "Ducheng Tan"
date: "2026-05-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE) 
```


```{r}
#install.packages("tidyverse")
#install.packages("palmerpenguins")
#install.packages("MASS")
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
library(ggplot2)
library(MASS)
library(ggplot2)
```


```{r}
palmerpenguins::penguins
glimpse(penguins)
```


```{r}
summary(penguins)
```
#ggplot 


```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = penguins$species)
) + geom_point(mapping = aes(color = species)) +  geom_smooth(method = "lm")
```
 
```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = species, shape =species)) +
  geom_smooth(method = "lm")
```
 
 
```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(aes(color = species, shape = species)) +
  geom_smooth(method = "lm") +
  labs(
    title = "Body mass and flipper length",
    subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
    x = "Flipper length (mm)", y = "Body mass (g)",
    color = "Species", shape = "Species"
  ) +
  scale_color_colorblind()
```
 
 
 1.2.5 Exercise 
 
How many rows are in penguins? How many columns?
344 x 8 (n x p) 

```{r}
nrow(penguins); ncol(penguins)
```


What does the bill_depth_mm variable in the penguins data frame describe? Read the help for ?penguins to find out.

a number denoting bill length (millimeters)  
```{r}
?penguins
```


Make a scatterplot of bill_depth_mm vs. bill_length_mm. That is, make a scatterplot with bill_depth_mm on the y-axis and bill_length_mm on the x-axis. Describe the relationship between these two variables.
There is a weak positive relation between the two variables. 

```{r}
ggplot(data = penguins, mapping = aes(x=bill_length_mm, y=bill_depth_mm, color = species)) + geom_point()
```


What happens if you make a scatterplot of species vs. bill_depth_mm? What might be a better choice of geom?
```{r}
ggplot(data = penguins, mapping= (aes(x= species,y= bill_depth_mm, color = species ))) +geom_boxplot()
```



Why does the following give an error and how would you fix it?
fixed via adding `mapping = aes(x, y)` 

```{r}
ggplot(data = penguins , mapping = aes(x= body_mass_g, y=flipper_length_mm, color =species)) +
  geom_point()
```



What does the na.rm argument do in geom_point()? What is the default value of the argument? Create a scatterplot where you successfully use this argument set to TRUE.

```{r}
ggplot(data = penguins , mapping = aes(x= body_mass_g, y=flipper_length_mm, color =bill_depth_mm)) +
  geom_point(na.rm = TRUE) # removes the missing data points without warning
```


Add the following caption to the plot you made in the previous exercise: “Data come from the palmerpenguins package.” Hint: Take a look at the documentation for labs().

```{r}
ggplot(data = penguins , mapping = aes(x= body_mass_g, y=flipper_length_mm, color =species)) +
  geom_point(na.rm = TRUE) +  labs(title = "body mass vs flipper_length",caption = "Data come from the palmerpenguins package.")
 # removes the missing data points without warning
```

Recreate the following visualization. What aesthetic should `bill_depth_mm` be mapped to? And should it be mapped at the global level or at the geom level?

```{r}
ggplot(data = penguins , mapping = aes(x=flipper_length_mm , y=body_mass_g, color =bill_depth_mm)) +
  geom_point(na.rm = TRUE, size = 3) + geom_smooth(method = "loess")
```


Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.
```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = island)
) +
  geom_point() +
  geom_smooth(se = FALSE)
```



Will these two graphs look different? Why/why not? 

```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point() +
  geom_smooth()

ggplot() +
  geom_point(
    data = penguins,
    mapping = aes(x = flipper_length_mm, y = body_mass_g)
  ) +
  geom_smooth(
    data = penguins,
    mapping = aes(x = flipper_length_mm, y = body_mass_g)
  )
```
 
 